### 2012-02-29 proposal/proposal.tex
added proposal document (empty)

\documentclass{ubicomp2011}
\usepackage{times}
\usepackage{url}
\usepackage{graphics}
\usepackage{color}
\usepackage[pdftex]{hyperref}
\hypersetup{%
pdftitle={Your Title}, pdfauthor={Your Authors}, pdfkeywords={your
keywords}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
breaklinks=true, }
\newcommand{\comment}[1]{}
\definecolor{Orange}{rgb}{1,0.5,0}
\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[#1]]}}}}

\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy

\begin{document}
% to make various LaTeX processors do the right thing with page size
\special{papersize=8.5in,11in}
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% use this command to override the default ACM copyright statement
% (e.g. for preprints). Remove for camera ready copy.

%\toappear{Paper submitted for evaluation in the Pervasive Computing Course, Spring 2012. The IT University of Copenhagen. Copyright remains with the authors.}

\toappear{Project proposal for the Pervasive Computing Course, Spring 2012. The IT University of Copenhagen. Copyright remains with the authors.}

\title{UbiComp 2011 Conference Proceedings Format}
\numberofauthors{2}
\author{
  \alignauthor Egil Hansen\\
    \affaddr{IT University of Copenhagen}\\
%    \affaddr{Address  (Blank for Blind Review)}\\
    \email{ekri@itu.dk}
 \alignauthor David Thomas\\
    \affaddr{IT University of Copenhagen}\\
%    \affaddr{Address  (Blank for Blind Review)}\\
    \email{dtho@itu.dk}  }
\maketitle

%\begin{abstract}
%  In this paper we describe the formatting requirements for the UbiComp 2011
%  Conference Proceedings, and offer recommendations on writing for the
%  worldwide Ubiquitous Computing readership.  Please review this document even if
%  you have submitted to Ubiquitous Computing conferences before, for some format
%  details have changed relative to previous years. These include the
%  formatting of table captions, the formatting of references, a
%  requirement to include ACM DL indexing information, and guidelines for how to handle relevant references to your own work while preparing your submission for blind review.
%\end{abstract}

%\keywords{Guides, instructions, authors' kit, conference publications.}

%\category{H.5.2}{Information interfaces and presentation (e.g., HCI)}{Miscellaneous}.

%\generalterms{This section is limited to the following 16 terms and
%MUST be included on the first page of all submissions after the ACM
%Categories section, then as well chosen properly on the Proceedings
%or Publication's submission page: Algorithms, Design,
%Documentation, Economics, Experimentation, Human Factors, Languages,
%Legal Aspects, Management, Measurement, Performance, Reliability,
%Security, Standardization, Theory, Verification.}

\section{Background}
hep

\section{Idea}
hep

\section{Scenario}
hep

\section{Plan}
hep

\subsection{Equipment}
hep

% Referencer:
% ~\cite{ref}

% til sidst
%\nocite{example-journal,example-abstracts,example-conference2}

\bibliographystyle{abbrv}
\bibliography{proposal}

\end{document}