### 2012-04-22 report/paper.tex
report, etc

\documentclass{ubicomp2011}
\usepackage{times}
\usepackage{url}
\usepackage{graphics}
\usepackage{color}
\usepackage[pdftex]{hyperref}
\hypersetup{%
pdftitle={Your Title}, pdfauthor={Your Authors}, pdfkeywords={your
keywords}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
breaklinks=true, }
\newcommand{\comment}[1]{}
\definecolor{Orange}{rgb}{1,0.5,0}
\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[#1]]}}}}

\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy

\begin{document}
% to make various LaTeX processors do the right thing with page size
\special{papersize=8.5in,11in}
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% use this command to override the default ACM copyright statement
% (e.g. for preprints). Remove for camera ready copy.
\toappear{Paper submitted for evaluation in the Pervasive Computing Course, Spring
2012. The IT University of Copenhagen.}

\title{Thingies for Dummies}
\numberofauthors{2}
\author{
  \alignauthor David Thomas\\
    \affaddr{Author}\\
    \affaddr{IT University of Copenhagen}\\
    \email{dtho@itu.dk}
 \alignauthor Egil Kristoffer Gorm Hansen\\
    \affaddr{Author}\\
    \affaddr{IT University of Copenhagen}\\
    \email{ekri@itu.dk}  }
\maketitle

\begin{abstract}
  TODO
\end{abstract}

\keywords{TDOD}

\category{TODO}{TODO}{TODO}.

\section{Introduction}

\section{Background/Related Work}
Seven problems with Smart Homes (ref)

\section{Thinges for Dummies Architecture}

\subsection{}

\section{Evaluation}

\section{Discussion}

\section{Future Work}

\section{Conclusion}

\section{Acknowledgements}

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}